---
title: "Thera Bank"
author: "Omar Ossama"
date: "6/8/2020"
output:
  pdf_document: default
  html_document: default
---

```{r include=FALSE}
setwd ("C:/Users/omaro/Documents/DSBA/03-Machine Learning/05 - Project")
```



Libraries used
```{r}
library(readxl)
library(esquisse)
library(corrplot)
library(ggplot2)
library(gridExtra)
library(DataExplorer)
library(cluster)
library(mice)
library(NbClust)
library(caTools)
library(rpart)
library(rattle)
library(rpart.plot)
library(randomForest)
library(ROCR)
library(ineq)
library(InformationValue)
library(caret)
```





```{r}
thera_bank = read_excel("Thera Bank_Personal_Loan_Modelling-dataset-1.xlsx", sheet="Bank_Personal_Loan_Modelling")
```



Exploratory data analysis
```{r}
head(thera_bank)
tail(thera_bank)
dim(thera_bank)
colnames(thera_bank)
str(thera_bank)
```



Editing the data
```{r}
colnames(thera_bank) = make.names(colnames(thera_bank))
sum(is.na(thera_bank))
md.pattern(thera_bank)
init.impute = mice(thera_bank, m=2, method = "pmm", seed = 1000)
thera_bank = complete(init.impute, 2)

##thera_bank$Family.members = [is.na(thera_bank$Family.members)] = 0

thera_bank = thera_bank [,-1]

thera_bank$Experience..in.years. = ifelse (thera_bank$Experience..in.years. < 0, -1*thera_bank$Experience..in.years., thera_bank$Experience..in.years.)

thera_bank$Personal.Loan = as.factor(thera_bank$Personal.Loan)
thera_bank$ZIP.Code = as.factor(thera_bank$ZIP.Code)
thera_bank$Education = as.factor(thera_bank$Education)
thera_bank$Securities.Account = as.factor(thera_bank$Securities.Account)
thera_bank$CD.Account = as.factor(thera_bank$CD.Account)
thera_bank$Online = as.factor(thera_bank$Online)
thera_bank$CreditCard = as.factor(thera_bank$CreditCard)

str(thera_bank)
```



```{r}
summary(thera_bank)
```


Univariate analysis
Checking the ratio of people who accepted the loan to those who did not
```{r}
thera_bank$Personal.Loan = as.numeric(thera_bank$Personal.Loan)
prop.table(table(thera_bank$Personal.Loan))
thera_bank$Personal.Loan = as.factor(thera_bank$Personal.Loan)
```



```{r}
plot_histogram_n_boxplot = function(variable, variableNameString, binw){
  h = ggplot(data = thera_bank, aes(x= variable))+
    labs(x = variableNameString,y ='count')+
    geom_histogram(fill = 'green',col = 'white',binwidth = binw)+
    geom_vline(aes(xintercept=mean(variable)),
            color="black", linetype="dashed", size=0.5)
  b = ggplot(data = thera_bank, aes('',variable))+ 
    geom_boxplot(outlier.colour = 'red',col = 'red',outlier.shape = 19)+
    labs(x = '',y = variableNameString)+ coord_flip()
  grid.arrange(h,b,ncol = 2)
}
```



```{r}
#Age in years
plot_histogram_n_boxplot(thera_bank$Age..in.years., 'Age', 1)

#Experience in years
plot_histogram_n_boxplot(thera_bank$Experience..in.years., 'Experience', 1)

#Income in K/month
plot_histogram_n_boxplot(thera_bank$Income..in.K.month., 'Income', 1)

#Family members
plot_histogram_n_boxplot(thera_bank$Family.members, 'Family members', 1)

#Credit Card Avg.
plot_histogram_n_boxplot(thera_bank$CCAvg, 'Credit Card Avg', 1)

#Mortgage
ggplot(thera_bank) +
 aes(x = Mortgage) +
 geom_histogram(bins = 30L, fill = "#6dcd59") +
 theme_minimal()

ggplot(thera_bank) +
 aes(x = "", y = Mortgage) +
 geom_boxplot(fill = "#6dcd59") +
 theme_minimal()

#Zip Code
ggplot(thera_bank) +
 aes(x = ZIP.Code) +
 geom_bar(fill = "#6dcd59") +
 theme_minimal()

#Personal loan
ggplot(thera_bank) +
 aes(x = Personal.Loan) +
 geom_bar(fill = "#6dcd59") +
 theme_minimal()

#Education
ggplot(thera_bank) +
 aes(x = Education) +
 geom_bar(fill = "#6dcd59") +
 theme_minimal()

#Securities
ggplot(thera_bank) +
 aes(x = Securities.Account) +
 geom_bar(fill = "#6dcd59") +
 theme_minimal()

#Credit Card Account
ggplot(thera_bank) +
 aes(x = CD.Account) +
 geom_bar(fill = "#6dcd59") +
 theme_minimal()

#Online use
ggplot(thera_bank) +
 aes(x = Online) +
 geom_bar(fill = "#6dcd59") +
 theme_minimal()

#Credit card
ggplot(thera_bank) +
 aes(x = CreditCard) +
 geom_bar(fill = "#6dcd59") +
 theme_minimal()
```



Bivariate analysis to see to effect of the different independant variable on the dependant variable (Personal Loan)
```{r}
ggplot(thera_bank, aes(fill = Personal.Loan, x = Age..in.years.)) + 
    geom_bar(position="fill")

ggplot(thera_bank, aes(fill = Personal.Loan, x = Experience..in.years.)) + 
    geom_bar(position="fill")

ggplot(thera_bank, aes(fill = Personal.Loan, x = Income..in.K.month.)) + 
    geom_bar(position="fill")

chisq.test(thera_bank$Personal.Loan, thera_bank$ZIP.Code)
ggplot(thera_bank, aes(fill = Personal.Loan, x = ZIP.Code)) + 
    geom_bar(position="fill")

ggplot(thera_bank, aes(fill = Personal.Loan, x = Family.members)) + 
    geom_bar(position="fill")

ggplot(thera_bank, aes(fill = Personal.Loan, x = CCAvg)) + 
    geom_bar(position="fill")

chisq.test(thera_bank$Personal.Loan, thera_bank$Education)
ggplot(thera_bank, aes(fill = Personal.Loan, x = Education)) + 
    geom_bar(position="fill")

ggplot(thera_bank, aes(fill = Personal.Loan, x = Mortgage)) + 
    geom_bar(position="fill")

chisq.test(thera_bank$Personal.Loan, thera_bank$Securities.Account)
ggplot(thera_bank, aes(fill = Personal.Loan, x = Securities.Account)) + 
    geom_bar(position="fill")

chisq.test(thera_bank$Personal.Loan, thera_bank$CD.Account)
ggplot(thera_bank, aes(fill = Personal.Loan, x = CD.Account)) + 
    geom_bar(position="fill")

chisq.test(thera_bank$Personal.Loan, thera_bank$Online)
ggplot(thera_bank, aes(fill = Personal.Loan, x = Online)) + 
    geom_bar(position="fill")

chisq.test(thera_bank$Personal.Loan, thera_bank$CreditCard)
ggplot(thera_bank, aes(fill = Personal.Loan, x = CreditCard)) + 
    geom_bar(position="fill")
```



```{r}
corrplot(cor(thera_bank[c(1:3,5,6,8)]),type="lower",method="number")
```


```{r}
chisq.test(thera_bank$CD.Account, thera_bank$ZIP.Code)
ggplot(thera_bank, aes(fill = CD.Account, x = ZIP.Code)) + 
    geom_bar(position="fill")

chisq.test(thera_bank$CD.Account, thera_bank$Education)
ggplot(thera_bank, aes(fill = CD.Account, x = Education)) + 
    geom_bar(position="fill")

chisq.test(thera_bank$CD.Account, thera_bank$Securities.Account)
ggplot(thera_bank, aes(fill = CD.Account, x = Securities.Account)) + 
    geom_bar(position="fill")

chisq.test(thera_bank$CD.Account, thera_bank$Online)
ggplot(thera_bank, aes(fill = CD.Account, x = Online)) + 
    geom_bar(position="fill")

chisq.test(thera_bank$CD.Account, thera_bank$CreditCard)
ggplot(thera_bank, aes(fill = CD.Account, x = CreditCard)) + 
    geom_bar(position="fill")
```




Adjusting variable types
```{r}
thera_bank$ZIP.Code = as.numeric(thera_bank$ZIP.Code)
thera_bank$Education = as.numeric(thera_bank$Education)
thera_bank$Personal.Loan = as.numeric(thera_bank$Personal.Loan)
thera_bank$Securities.Account = as.numeric(thera_bank$Securities.Account)
thera_bank$CD.Account = as.numeric(thera_bank$CD.Account)
thera_bank$Online = as.numeric(thera_bank$Online)
thera_bank$CreditCard = as.numeric(thera_bank$CreditCard)
str(thera_bank)
```



Clustering using K means method
```{r}
seed = 1000
set.seed(seed)

thera_bank_scaled = scale(thera_bank)

totWss = rep (0,5)
for (k in 1:5){
  set.seed(seed)
  clust = kmeans(x=thera_bank_scaled, centers=k, nstart=5)
  totWss[k] = clust$tot.withinss
}
print(totWss)
plot(c(1:5),totWss,type="b")

nc = NbClust(thera_bank_scaled, min.nc=2, max.nc = 5, method = "kmeans")


```
We can see that the optimum number of clusters is = 3

Creating 3 clusters
```{r}
clust = kmeans(x=thera_bank_scaled, centers=3, nstart=5)
clusplot(thera_bank_scaled,clust$cluster,color=TRUE,shade=TRUE)

thera_bank_clust = thera_bank
thera_bank_clust$Cluster = clust$cluster
print(thera_bank_clust)

cust_profile = aggregate(thera_bank_scaled, list(thera_bank_clust$Cluster), FUN="mean")
print(cust_profile)
```


Building CART and random forest models
Splitting the data
```{r}
summary(thera_bank)
thera_bank$Personal.Loan = as.factor(thera_bank$Personal.Loan)

set.seed(seed)

sample = sample.split(thera_bank$Personal.Loan,SplitRatio = 0.7)
train = subset(thera_bank,sample == TRUE)
test = subset(thera_bank,sample == FALSE)

nrow(train)
nrow(test)

prop.table(table(thera_bank$Personal.Loan))
prop.table(table(train$Personal.Loan))
prop.table(table(test$Personal.Loan))

```



CART model
```{r}
r.ctrl = rpart.control(minsplit = 10, minbucket = 3, xval = 10)

thera_cart_model <- rpart(formula = Personal.Loan~., data = train, method = "class", control = r.ctrl)
thera_cart_model
fancyRpartPlot(thera_cart_model)
prp(thera_cart_model)
```


Variable importance
```{r}
thera_cart_model$variable.importance
```


```{r}
thera_cart_model$cptable
```
No pruning required



Model validation using the train data
```{r}
train_class = predict(thera_cart_model, train, type="class")
train_score = predict(thera_cart_model, train)


caret::confusionMatrix(train$Personal.Loan, train_class)
```
Accuracy = 98.57%

Model evaluation using test data
```{r}
test_class = predict(thera_cart_model, test, type="class")
test_score = predict(thera_cart_model, test)

caret::confusionMatrix(test$Personal.Loan, test_class)
```
Accuracy = 97.87%
Difference between train and test accuracy = 0.7%
So we can say that the model is valid




```{r}
set.seed(seed)

thera_rf = tuneRF(x = train[, -9],
                  y = train$Personal.Loan,
                  mtrystart = 3,
                  stepfactor=1.5,
                  ntree=70,
                  improve=0.0001,
                  nodesize=10,
                  trace=TRUE,
                  plot=TRUE,
                  doBest=TRUE,
                  importance=TRUE 
                  )

print(thera_rf)

plot(thera_rf)

print(thera_rf$importance)
```
Optimum mtry value = 6
OOB = 1.26%


Model validation
```{r}
train_class_rf = predict(thera_rf, train, type="class")
train_score_rf = predict(thera_rf, train)

caret::confusionMatrix(train$Personal.Loan, train_class_rf)
```
Accuracy = 99.34%



```{r}
test_class_rf = predict(thera_rf, test, type="class")
test_score_rf = predict(thera_rf, test)

caret::confusionMatrix(test$Personal.Loan, test_class_rf)
```
Accuracy = 98.13%


Model performance measures
```{r}
train_cart = train
test_cart = test

train_rf = train
test_rf = test


train_cart$prediction = predict(thera_cart_model, train_cart, type="class")
train_cart$prob1 = predict(thera_cart_model, train_cart, type="prob")[,"1"]
summary(train_cart)

test_cart$prediction = predict(thera_cart_model, test_cart, type="class")
test_cart$prob1 = predict(thera_cart_model, test_cart, type="prob")[,"1"]
head(test_cart)

train_rf$prediction = predict(thera_rf, train_rf, type="class")
train_rf$prob1 = predict(thera_rf, train_rf, type="prob")[,"1"]
head(train_rf)

test_rf$prediction = predict(thera_rf, test_rf, type="class")
test_rf$prob1 = predict(thera_rf, test_rf, type="prob")[,"1"]
head(test_rf)

summary(test_rf)
```




Performance metrics for CART train
```{r}

pred_train_cart = prediction(train_cart$prob1, train_cart$Personal.Loan)
perf_train_cart = performance(pred_train_cart, "tpr", "fpr")

plot(perf_train_cart)

KS_train_cart = max(perf_train_cart@y.values[[1]]-perf_train_cart@x.values[[1]])
print(KS_train_cart)

auc_train_cart = performance(pred_train_cart,"auc"); 
auc_train_cart = as.numeric(auc_train_cart@y.values)
print(auc_train_cart)

gini_train_cart = ineq(train_cart$prob1, type="Gini")
print(gini_train_cart)
```
KS value = 0.9126
AUC = 0.9815
Gini = 0.8705



Performance metrics for CART test
```{r}

pred_test_cart = prediction(test_cart$prob1, test_cart$Personal.Loan)
perf_test_cart = performance(pred_test_cart, "tpr", "fpr")

plot(perf_test_cart)

KS_test_cart = max(perf_test_cart@y.values[[1]]-perf_test_cart@x.values[[1]])
print(KS_test_cart)

auc_test_cart = performance(pred_test_cart,"auc"); 
auc_test_cart = as.numeric(auc_test_cart@y.values)
print(auc_test_cart)

gini_test_cart = ineq(test_cart$prob1, type="Gini")
print(gini_test_cart)
```
KS value = 0.9243
AUC = 0.9826
Gini = 0.8739



Performance metrics for random forest train
```{r}

pred_train_rf = prediction(train_rf$prob1, train_rf$Personal.Loan)
perf_train_rf = performance(pred_train_rf, "tpr", "fpr")

plot(perf_train_rf)

KS_train_rf = max(perf_train_rf@y.values[[1]]-perf_train_rf@x.values[[1]])
print(KS_train_rf)

auc_train_rf = performance(pred_train_rf,"auc"); 
auc_train_rf = as.numeric(auc_train_rf@y.values)
print(auc_train_rf)

gini_train_rf = ineq(train_rf$prob1, type="Gini")
print(gini_train_rf)
```
KS = 0.9943
AUC = 0.9999
Gini = 0.8987



Performance metrics for random forest test
```{r}

pred_test_rf = prediction(test_rf$prob1, test_rf$Personal.Loan)
perf_test_rf = performance(pred_test_rf, "tpr", "fpr")

plot(perf_test_rf)

KS_test_rf = max(perf_test_rf@y.values[[1]]-perf_test_rf@x.values[[1]])
print(KS_test_rf)

auc_test_rf = performance(pred_test_rf,"auc"); 
auc_test_rf = as.numeric(auc_test_rf@y.values)
print(auc_test_rf)

gini_test_rf = ineq(test_rf$prob1, type="Gini")
print(gini_test_rf)
```
KS = 0.9662
AUC = 0.9979
Gini = 0.8991





Concordance and discordcance ratios:
```{r}
conc_train_cart = Concordance(actuals=train_cart$Personal.Loan, predictedScores=train_cart$prob1)
print(conc_train_cart)

conc_test_cart = Concordance(actuals=test_cart$Personal.Loan, predictedScores=test_cart$prob1)
print(conc_test_cart)

conc_train_rf = Concordance(actuals=train_rf$Personal.Loan, predictedScores=train_rf$prob1)
print(conc_train_rf)

conc_test_rf = Concordance(actuals=test_rf$Personal.Loan, predictedScores=test_rf$prob1)
print(conc_test_rf)

```

